---
title: "Prevalence Analysis"
author: "Akacia Halliday-Isaac"
date: "April 9, 2020"
output:
  html_document: default
  word_document: default
---

Read in data
```{r}
population <- read.csv("population.csv",header=TRUE)
habitat <- read.csv("habitatcomp.csv", header = TRUE)
```


Install Packages
```{r}
library(ggplot2)
library(coin)
library(pwr)
library(plyr)
library(dplyr)
library(survival)
library(lme4)
library(Matrix)
library(plotly)
library(multcompView)
library(visreg)
library(jtools)
library(webshot)
library(agricolae)
library(knitr)
library('dplyr')      # for data manipulation
library('tidyr')      # for reshaping data

library('ggplot2')    # plotting data
library('scales')     # for scale_y_continuous(label = percent)
library('ggthemes')   # for scale_fill_few('medium')

library('ztable')     # format tables for reporting
```


Stacked bar graph of sites
```{r}
p <- plot_ly(data=population, x = ~Site, y = ~Infected, type = 'bar', textposition = 'outside', name = 'Infected') %>%
    add_trace(y = ~Uninfected, name = 'Uninfected') %>%
    layout(barmode = 'stack', xaxis = list(tickangle = 45, title = "Site", titlefont=list(size=15)),
 yaxis = list(title = " # of S. adustus", titlefont=list(size=15), range = c(0, 20))) 
p
```



 Determine if there is a difference in infection prevalence between sites
```{r}
site <- read.csv("siteanalysis.csv", header = T)
site
```
 
```{r}
dm <-data.matrix(population[4:5])
dimnames(dm) <- population[1]
```


```{r}
fisher.test(dm, simulate.p.value = TRUE)
```

```{r}
library(rcompanion)

PT = pairwiseNominalIndependence(dm,

                                 fisher = TRUE,

                                 gtest  = FALSE,

                                 chisq  = FALSE,
                                 digits = 3)

PT
```

```{r}
library(rcompanion)

letters <- cldList(comparison = PT$Comparison,
        p.value    = PT$p.adj.Fisher,
        threshold  = 0.05)

letters
```

```{r}
plot_ly(data=population, x = ~Site, y = ~Infected, type = 'bar',  text = ~letters$Letter, textposition = 'outside', name = 'Infected') %>%
    add_trace(y = ~Uninfected, name = 'Uninfected') %>%
    layout(barmode = 'stack', xaxis = list(tickangle = 45, title = "Site", titlefont=list(size=15)),
 yaxis = list(title = " Number of S. adustus Sampled", titlefont=list(size=15),  range = c(0, 30))) 

```





Determine if there is a difference in fish populations between sites
 
```{r}
dm2 <-data.matrix(population[11:12])
dimnames(dm2) <- population[1]
```


```{r}
fisher.test(dm2, simulate.p.value = TRUE)
```

Calculating Habitat complexity using Shannon-Weiner Diversity Index
```{r}
library(vegan)
diversity(habitat[-1], index="shannon")
```

Attaching Habitat Complexity to datasheet
```{r}

 population$HabitatComplexity <- diversity(habitat[-1], index="shannon") 
 population$HabitatComplexity
```



Mixed model
```{r}
library(lmerTest)
m0 <- lmer(Prevalence ~ Population.Density+Stegastes.Population.Density+HabitatComplexity+Average.Zonal.Velocity+Average.Meridional.Velocity+NEAR_DIST + Population.Density:Stegastes.Population.Density +  + Population.Density:Average.Zonal.Velocity + Population.Density:Average.Meridional.Velocity + Population.Density:NEAR_DIST + Average.Meridional.Velocity:Average.Zonal.Velocity + Population.Density:NEAR_DIST + (1|Season), data=population)
```

```{r}
anova(m0)
```

```{r}
summary(m0)
```

Multiple Regression

Create a subset that removes NA's
```{r}
subset <- subset(population, Population.Density!='NA' & Stegastes.Population.Density!='NA' & HabitatComplexity !='NA' & Average.Zonal.Velocity!='NA' & Average.Meridional.Velocity!='NA' & NEAR_DIST!='NA')
View(subset) # do this to check that it looks OK, then close it
dim(subset)
```

Fit model that only includes main effects
```{r}
fit1 <- lm(Prevalence ~ Population.Density+Stegastes.Population.Density+HabitatComplexity+Average.Zonal.Velocity+Average.Meridional.Velocity+NEAR_DIST, data=subset, na.action = "na.fail")
library(car)
vif(fit1) # get VIF for each predictor
mean(vif(fit1))
```

Fit model that includes desired interactions
```{r}
fit2 <- lm(Prevalence ~ Population.Density+Stegastes.Population.Density+HabitatComplexity+Average.Zonal.Velocity+Average.Meridional.Velocity+NEAR_DIST+Average.Zonal.Velocity:Average.Meridional.Velocity+Average.Meridional.Velocity:NEAR_DIST+Population.Density:Stegastes.Population.Density + Population.Density:HabitatComplexity,  data=subset, na.action = "na.fail")
# check assumptions and transform variables if necessary
hist(fit2$residuals)
plot(fit2$residuals ~ fit2$fitted.values)
```

Transform Data & Recheck Assumptions
```{r}
subset$sqrt_Prevalence <- sqrt(subset$Prevalence)
fit3 <- lm(sqrt_Prevalence ~ Population.Density+Stegastes.Population.Density+HabitatComplexity+Average.Zonal.Velocity+Average.Meridional.Velocity+NEAR_DIST+Average.Zonal.Velocity:Average.Meridional.Velocity+Average.Meridional.Velocity:NEAR_DIST+Population.Density:Stegastes.Population.Density+Population.Density:HabitatComplexity, data=subset, na.action = "na.fail")
# check assumptions and transform variables if necessary
hist(fit3$residuals)
plot(fit3$residuals ~ fit3$fitted.values)
```

```{r}
summary.lm(fit3)
```

```{r}
library(MuMIn)
dd <- dredge(fit3)
```

Show table of all models within 10 AICc units of the best model:
```{r}
AICc <- subset(dd, delta < 10)
write.csv(AICc, "aic.csv")
```

since several models are within 2 AICc units of the best model, use model averaging, and
calculate relative variable importance (RVI) for each predictor:
```{r}
importance(dd)
```

Fit best model
```{r}
fit4 <- lm(sqrt_Prevalence ~ Population.Density+Average.Meridional.Velocity+NEAR_DIST, data=subset, na.action = "na.fail")
Anova(fit4, type='II')
```

```{r}
summary.lm(fit4)
```

Illustrate effect of Average.Meridional.Velocity on prevalence
```{r}
g2 <- ggplot(subset,aes(y=sqrt_Prevalence,x=Average.Meridional.Velocity))+geom_point()+ geom_smooth(method = lm)+ labs(x = "Avg. Meridional Velocity (sverdrup)", y = "Prevelence (proportion infected per pop.)")

g2 + annotate(geom="text", x=0.03, y=1, label="p= 0.00318")+ theme(panel.background = element_blank())

ggsave("amv v prev.png")
```

Illustrate effects of Population Density on Prevalence
```{r}
library(ggplot2)
g <- ggplot(subset,aes(y=sqrt_Prevalence,x=Population.Density))+geom_point()+ geom_smooth(method = lm)+ labs(x = expression(paste(italic("S. adustus "), "Population Density (# of individuals)")), y = "Prevelence (proportion infected per pop.)")
g + annotate(geom="text", x=8, y=0.9, label="p= 0.04771")+ theme(panel.background = element_blank())

ggsave("pop den v prev.png")
```


Multivariate analysis of habitat variables
```{r}
library(vegan)
```

Read in data
```{r}
env <- read.csv("habitatcomp2.csv", header = TRUE)
```

Bray-Curtis distances between samples
```{r}
dis <- vegdist(env)
dis
```

Calculate multivariate dispersions
```{r}
mod <- betadisper(dis, group = habitat$Site)
mod
```

Load factoextra for visualization
```{r}
library(factoextra)
```

Compute PCA
```{r}
res.pca <- prcomp(env, scale = TRUE)
```

Visualize eigenvalues (scree plot). Show the percentage of variances explained by each principal component.
```{r}
fviz_eig(res.pca)
```


Graph of individuals. Individuals with a similar profile are grouped together.
```{r}
fviz_pca_ind(res.pca,
             col.ind = "cos2", # Color by the quality of representation
             gradient.cols = c("#00AFBB", "#E7B800", "#FC4E07"),
             repel = TRUE     # Avoid text overlapping
             )
```




Graph of variables. Positive correlated variables point to the same side of the plot. Negative correlated variables point to opposite sides of the graph.

```{r}
fviz_pca_var(res.pca,
             col.var = "contrib", # Color by contributions to the PC
             gradient.cols = c("#00AFBB", "#E7B800", "#FC4E07"),
             repel = TRUE     # Avoid text overlapping
             )
```

Biplot of individuals and variables
```{r}
fviz_pca_biplot(res.pca, repel = TRUE,
                col.var = "#2E9FDF", # Variables color
                col.ind = "#696969"  # Individuals color
                )
```

Results
```{r}
library(factoextra)
# Eigenvalues
eig.val <- get_eigenvalue(res.pca)
eig.val
  
# Results for Variables
res.var <- get_pca_var(res.pca)
res.var$coord          # Coordinates
res.var$contrib        # Contributions to the PCs
res.var$cos2           # Quality of representation 
# Results for individuals
res.ind <- get_pca_ind(res.pca)
res.ind$coord          # Coordinates
res.ind$contrib        # Contributions to the PCs
res.ind$cos2           # Quality of representation
```


```{r}
groups <- population$Infection.Level
fviz_pca_ind(res.pca,
             col.ind = groups, # color by groups
             palette = c("#00AFBB",  "#FC4E07", "#71D1CC"),
             addEllipses = TRUE, # Concentration ellipses
             ellipse.type = "confidence",
             legend.title = "Infection Level",
             repel = TRUE
             )
dev.copy(png,file='PCA2.png', width = 5, 
         height = 5, units = 'in', res = 800)
#if this fails, then lower the resolution (res) 
dev.off() # necessary step
```

Look at relationship between Population Density and Distance of Populations from Watersheds
```{r}
cor.test(population$Population.Density, population$NEAR_DIST, method = "pearson")
```

```{r}
cor.test(population$Stegastes.Population.Density, population$NEAR_DIST, method =
            "pearson")
```


Stepwise Regression of habitat composition components and prevalence
```{r}
lmMod <- lm(Prevalence ~ X..Rock + X.Seagrass + X.Sandy.Bottom + X..Coral + X..Sponge + X..Rubble , data = population)
selectedMod <- step(lmMod)
summary(selectedMod)
```

```{r}
h <- lm(Prevalence ~ X.Sandy.Bottom, data = population)
summary.lm(h)
```

