---
title: "Length-Weight Relationships"
author: "Akacia Halliday"
date: "January 14, 2018"
output:
  word_document: default
  html_document: default
---


Install Packages
```{r}
library(FSA)
library(dplyr)
library(knitr)
```

Read in Data
```{r}
relationship <- read.csv("lengthweight3.csv",header=TRUE)
```

View data
```{r}
str(relationship)
```
Create summary table of data
```{r}
t <- relationship %>%
    group_by(Health) %>%
summarize(count = n(),
              MeanWeight = mean(Mass), MeanLength = mean(Length),
              StDevWeight = sd(Mass), StDevLength = sd(Length))
t
```


```{r kable, results = 'asis'}
kable(head(t), digits = 2)

```

Create a subset containing just the length and weight variables
```{r}
lengthweight <- Subset(relationship,!is.na(Mass) & !is.na(Length))
```

Create Log(Length) and Log(Weight) variable
```{r}
lengthweight$logL <- log(lengthweight$Length)
lengthweight$logW <- log(lengthweight$Mass)
```

View changes
```{r}
str(lengthweight)

```

Fit transformed model with Linear Model 
```{r}
lm1 <- lm(logW~logL,data=lengthweight)
```

View data and model fit
```{r}
fitPlot(lm1,xlab="log Total Length (mm)",ylab="log Weight (g)",main="Linear Model of S. adustus")
```

Summary of model fit
```{r}
summary(lm1)
```

A test of whether the fish in a population exhibit isometric growth or not can be obtained by noting that
b is the estimated slope from fitting the transformed length-weight model. The slope is generically labeled
with ?? such that the test for allometry can be translated into the following statistical hypotheses:
? H0 : B = 3  H0 : "Isometric growth"
? HA : B not = 3 HA : "Allometric growth"


Determine whether damselfish are exhibiting allometric growth
```{r}
hoCoef(lm1,2,3)

confint(lm1)
```

Correct back transformation bias
```{r}
syx <- summary(lm1)$sigma
( cf <- exp((syx^2)/2) ) # this is the bias correction factor
```

Predict mean log weight
```{r}
( pred.log <- predict(lm1,data.frame(logL=log(100)),interval="c") )
```

Use predicted mean log weight to form a biased predicted mean weight
```{r}
( bias.pred.orig <- exp(pred.log) ) # biased prediction on original scale
```

Compute bias-corrected back-transformed mean weight
```{r}
( pred.orig <- cf*bias.pred.orig ) # corrected prediction on original scale
```

Compare Length-Weight Relationships between groups via model
```{r}
lengthweight$Health <- as.factor(lengthweight$Health)
lm2 <- lm(logW~logL*Health,data=lengthweight)
summary(lm2)
```

Analyze model using ANOVA
```{r}
anova(lm2)

```

Plot comparison model
```{r}
colors <- c("#56B4E9", "#E69F00")

fitPlot(lm2,xlab="Fork Length (mm)",ylab="Mass (g)",legend="topleft", col=c('red', 'green', 'blue'))

dev.copy(png,file='LM v health.png', width = 6, 
         height = 6, units = 'in', res = 800)
#if this fails, then lower the resolution (res) 
dev.off()

```

```{r}
fitPlot(lm2,xlab="Fork Length (mm)",ylab="Mass (g)",legend="topleft",main="S. adustus Length-Mass Relationship by Health Status")

dev.copy(png,file='LM v health2.png', width = 6, 
         height = 6, units = 'in', res = 800)
#if this fails, then lower the resolution (res) 
dev.off()
```

Infection Quantification analysis
```{r}
q <- read.csv("Infection Quantification.csv",header=TRUE)
names(q)
```

Create a subset containing just the length and weight variables
```{r}
quant <- subset(q,!is.na(Weight..g.) & !is.na(TL..mm.))
```

Create Log(Length) and Log(Weight) variable
```{r}
quant$logL <- log(quant$TL..mm.)
quant$logW <- log(quant$Weight..g.)
```

Check for outliers in dataset 
```{r}
library(EnvStats)
test <- rosnerTest(quant$Infected.cell.per.RBC.load,
  k = 5
)
test
```

Remove the 2 identified outlier
```{r}
test <- quant[-c(54),] 
quant <- quant[-c(70),]
```

Compare Length-Weight Relationships by Infection level
```{r}
lm3 <- lm(logW~logL*Infected.cell.per.RBC.load,data=quant)
summary(lm3)
```

Analyze model using ANOVA
```{r}
anova(lm3)

```


3D scatterplot for Length by Mass by Infection Level 
```{r}
library(scatterplot3d)
scatterplot3d(quant$logL, quant$logW, quant$Infected.cell.per.RBC.load, pch = 16, grid = TRUE, box = FALSE, xlab = "Fork Length (mm)", ylab = "Mass (g)", zlab = "Infection Level")

dev.copy(png,file='LM v Infection lvl.png', width = 6, 
         height = 6, units = 'in', res = 800)
#if this fails, then lower the resolution (res) 
dev.off()
```

Compare Length-Weight Relationships by WBC Number
```{r}
lm4 <- lm(logW~logL*X..WBCS,data=quant)
summary(lm4)
```

Analyze model using ANOVA
```{r}
anova(lm4)

```


```{r}
library(scatterplot3d)
scatterplot3d(quant$logL, quant$logW, quant$X..WBCS, pch = 16, grid = TRUE, box = FALSE, xlab = "Fork Length (mm)", ylab = "Mass (g)", zlab = "Number of White Blood Cells")

dev.copy(png,file='LM v wbcs.png', width = 6, 
         height = 6, units = 'in', res = 800)
#if this fails, then lower the resolution (res) 
dev.off()
```

```{r}
library("ggpubr")
ggscatter(quant, x = "Infected.cell.per.RBC.load", y = "logL", 
          add = "reg.line", conf.int = TRUE, 
          cor.coef = TRUE, cor.method = "pearson", repel = TRUE, cor.coef.coord = c(0,4.65), 
          xlab = "Infection Level", ylab = "Fork Length (mm)")

dev.copy(png,file='Infection Level v Length.png', width = 6, 
         height = 6, units = 'in', res = 800)
#if this fails, then lower the resolution (res) 
dev.off()


```

```{r}
library("ggpubr")
ggscatter(quant, x = "Infected.cell.per.RBC.load", y = "logW", 
          add = "reg.line", conf.int = TRUE, 
          cor.coef = TRUE, cor.method = "pearson", repel = TRUE, cor.coef.coord = c(0,3),
          xlab = "Infection Level", ylab = "Mass (mg)")

dev.copy(png,file='Infection Level v Mass.png', width = 6, 
         height = 6, units = 'in', res = 800)
#if this fails, then lower the resolution (res) 
dev.off()
```



Does the Number of White Blood Cells present correlate to infection intensity
```{r}
cor.test(quant$X..WBCS, quant$Infected.cell.per.RBC.load, method =
            "pearson")

```



```{r}
library("ggpubr")
ggscatter(quant, x = "Infected.cell.per.RBC.load", y = "X..WBCS", 
          add = "reg.line", conf.int = TRUE, 
          cor.coef = TRUE, cor.method = "pearson", repel = TRUE, cor.coef.coord = c(0,375),
          xlab = "Infection Level", ylab = "Number of White Blood Cells Present")

dev.copy(png,file='Infection Level v wbcs.png', width = 6, 
         height = 6, units = 'in', res = 800)
#if this fails, then lower the resolution (res) 
dev.off()
```

Fit log model to White Blood Cells present vs infection intensity
```{r}
lm_log.model = lm(log1p(X..WBCS) ~ log1p(Infected.cell.per.RBC.load), data = quant)

```
Results of the model
```{r}
summary(lm_log.model)
```
```{r}
quant$logwbcs <- log1p(quant$X..WBCS)
```

```{r}
quant$logicprbc <- log1p(quant$Infected.cell.per.RBC.load)
```

```{r}
library(ggplot2)

ggplot(quant, aes(x = logwbcs, y = logicprbc)) + 
  geom_point() +
  stat_smooth(method = "lm", col = "red") +
  xlab("Number of White Blood Cells") + ylab("Infection Level")
```


```{r}
plot(X..WBCS~Infected.cell.per.RBC.load, data=quant)
```

```{r}
quant$X..WBCS <- as.numeric(gsub("\\.", "", quant$X..WBCS))
quant$Infected.cell.per.RBC.load <- as.numeric(gsub("\\.", "", quant$Infected.cell.per.RBC.load))
```



Look at uninfected data
```{r}
q2 <- read.csv("uninfected fish.csv",header=TRUE)
names(q2)
```

```{r}
plot(X..WBCS~X..Uninfected.Cells, data=q2)
```

Does the Number of White Blood Cells present correlate to cell count
```{r}
cor.test(q2$X..WBCS, q2$X..Uninfected.Cells, method =
            "pearson")

```


Fit log model to White Blood Cells present vs uninfected cell count
```{r}
lm_log.model2 = lm(log1p(X..WBCS) ~ log1p(X..Uninfected.Cells), data = q2)

```
Results of the model
```{r}
summary(lm_log.model2)
```

```{r}
q2$logwbcs <- log1p(q2$X..WBCS)
```

```{r}
q2$logrbc <- log1p(q2$X..Uninfected.Cells)
```

```{r}
library(ggplot2)

ggplot(q2, aes(x = logwbcs, y = logrbc)) + 
  geom_point() +
  stat_smooth(method = "lm", col = "red")+
  xlab("White Blood Cell Count") + ylab("Red Blood Cell Count")
```

```{r}
wbcs<- read.csv("wbcs.csv",header=TRUE)
View(wbcs)
```


```{r}
library(dplyr)
group_by(wbcs, Group) %>%
  summarise(
    count = n(),
    mean = mean(WBCS, na.rm = TRUE),
    sd = sd(WBCS, na.rm = TRUE)
  )
```


Compute the analysis of variance
```{r}
res.aov <- aov(WBCS ~ Group, data = wbcs)
summary(res.aov)
```

Visualize Results
```{r}
library("ggpubr")
my_comparisons <- list( c("Infected", "Zero Intensity"), c("Zero Intensity", "Uninfected"), c("Infected", "Uninfected") )
ggboxplot(wbcs, x = "Group", y = "WBCS", 
          color = "Group", palette = c("red", "blue", "green"),
          order = c("Infected", "Zero Intensity", "Uninfected"),
          ylab = "Leukocyte Count")+ 
  stat_compare_means(comparisons = my_comparisons)+ # Add pairwise comparisons p-value
  stat_compare_means(label.y = 50)     # Add global p-value

dev.copy(png,file='wbcs by group.png', width = 8, 
         height = 6, units = 'in', res = 800)
#if this fails, then lower the resolution (res) 
dev.off()
```



